---
title: "Determining the % of Core ASVs that are captured by swabs of different colors"
author: "Skarlupka"
date: '2024-05-17'
output: html_document
---

#Intro
```{r}
#The goal of this code is the pull out a list of ASVs that could be considered "Core", and then determining what percentage of those core ASVs are captured by the swabs.

```
#Set working directory
```{r}

setwd("INSERT ADDRESS HERE")

```

#Loading Libraries
```{r, Libraries}
library("ggplot2")
library("phyloseq")
library("vegan")
library("DESeq2")
library("reltools")
library("phylosmith")
library("gridExtra")

set.seed(81471438)
```

#Importing files of interest
```{r, import the files of interest}
#Using the rarefied and abundance cutoff phyloseq object from "Swab Photo Score Analysis.Rmd". 

physeq.swab1.abund

```


#Rumen Liquid - Determine which ASVs are Core to the Rumen Liquids
```{r}
#This chunk was generated with ChatGPT

# Subset the phyloseq object based on the metadata value
subset_physeq <- subset_samples(physeq.swab1.abund, PercentileName == "RL")

# Convert the abundance data to presence/absence data (binary data)
pa_physeq <- transform_sample_counts(subset_physeq, function(x) ifelse(x > 0, 1, 0))

# Calculate the prevalence of each ASV
prevalence <- apply(otu_table(pa_physeq), 2, function(x) sum(x) / length(x))

# Determine the ASVs present in more than 80% of samples
threshold <- 0.8
bac_RL_asv_ids <- names(prevalence[prevalence > threshold])

# Output the result
bac_RL_asv_ids
#242 ASVs identified

```


##Rumen Liquid - Remove ASVs that are not in the "Core" liquid list
```{r}
physeq.RLPrune <- prune_taxa(taxa_names(physeq.swab1.abund) %in% bac_RL_asv_ids , physeq.swab1.abund)

```


##Rumen Liquid - Count how many ASVs from the "Core Liquid" list are present
```{r}
#Convert counts to 1 if they are >0.
pa_physeq <- transform_sample_counts(physeq.RLPrune, function(x) ifelse(x > 0, 1, 0))

#Add a new column in the original metadata. This column will take the result of a sum of the ASV values for each sample divided by the length of the ASV table. 
sample_data(physeq.swab1.abund)$RLPrev <- rowSums(otu_table(pa_physeq)) / ntaxa(pa_physeq)


```

##Rumen Liquid - Plot the percentage of prevalent ASVs captured 
```{r}
physeq.swab1 <- subset_samples(physeq.swab1.abund, Swab_Rumen=="Swab")

b_RL_Prev <- ggplot(sample_data(physeq.swab1), aes(x = Darkness, y = RLPrev)) +
  geom_point() +  # Add points to the plot
  geom_smooth(method = "lm", se = FALSE) +  # Add a line of best fit
  labs(x = "Darkness Score", y = "% of Liquid Bacterial ASVs Captured") +  # Add axis labels and title
  theme_minimal()  # Optional: apply a minimal theme to the plot

b_RL_Prev


cor.test(sample_data(physeq.swab1)$Darkness, sample_data(physeq.swab1)$RLPrev, method = "spearman")
# data:  sample_data(physeq.swab1)$Darkness and sample_data(physeq.swab1)$RLPrev
# S = 5953053, p-value = 5.223e-07
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#      rho 
# 0.259389


cor.test(sample_data(physeq.swab1)$Darkness, sample_data(physeq.swab1)$RLPrev, method = "kendall")
# data:  sample_data(physeq.swab1)$Darkness and sample_data(physeq.swab1)$RLPrev
# z = 4.9364, p-value = 7.959e-07
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#      tau 
# 0.173783


```




#Rumen Solid - Determine which ASVs are Core to the Rumen Solids
```{r}
#Generated with ChatGPT

# Subset the phyloseq object based on the metadata value
subset_physeq <- subset_samples(physeq.swab1.abund, PercentileName == "RS")

# Convert the abundance data to presence/absence data (binary data)
pa_physeq <- transform_sample_counts(subset_physeq, function(x) ifelse(x > 0, 1, 0))

# Calculate the prevalence of each ASV
prevalence <- apply(otu_table(pa_physeq), 2, function(x) sum(x) / length(x))

# Determine the ASVs present in more than 80% of samples
threshold <- 0.8
bac_RS_asv_ids <- names(prevalence[prevalence > threshold])

# Output the result
bac_RS_asv_ids

#396 ASVs identified

```


##Rumen Solid - Remove ASVs that are not in the "Core" liquid list
```{r}
physeq.RSPrune <- prune_taxa(taxa_names(physeq.swab1.abund) %in% bac_RS_asv_ids , physeq.swab1.abund)

```


##Rumen Solid - Count how many ASVs from the "Core Liquid" list are present
```{r}
#Convert counts to 1 if they are >0.
pa_physeq <- transform_sample_counts(physeq.RSPrune, function(x) ifelse(x > 0, 1, 0))

#Add a new column in the metadata. This column will take the result of a sum of the ASV values for each sample divided by the length of the ASV table. 
sample_data(physeq.swab1.abund)$RSPrev <- rowSums(otu_table(pa_physeq)) / ntaxa(pa_physeq)


```

##Rumen Solid - Plot the percentage of prevalent ASVs captured 
```{r}
physeq.swab1 <- subset_samples(physeq.swab1.abund, Swab_Rumen=="Swab")

b_RS_Prev <- ggplot(sample_data(physeq.swab1), aes(x = Darkness, y = RSPrev)) +
  geom_point() +  # Add points to the plot
  geom_smooth(method = "lm", se = FALSE) +  # Add a line of best fit
  labs(x = "Darkness Score", y = "% of Solid Bacterial ASVs Captured") +  # Add axis labels and title
  theme_minimal()  # Optional: apply a minimal theme to the plot

b_RS_Prev


cor.test(sample_data(physeq.swab1)$Darkness, sample_data(physeq.swab1)$RSPrev, method = "spearman")
# data:  sample_data(physeq.swab1)$Darkness and sample_data(physeq.swab1)$RSPrev
# S = 5987143, p-value = 8.095e-07
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.2551479 

cor.test(sample_data(physeq.swab1)$Darkness, sample_data(physeq.swab1)$RSPrev, method = "kendall")
# data:  sample_data(physeq.swab1)$Darkness and sample_data(physeq.swab1)$RSPrev
# z = 4.794, p-value = 1.635e-06
# alternative hypothesis: true tau is not equal to 0
# sample estimates:
#       tau 
# 0.1685668 

grid.arrange(b_RS_Prev, b_RL_Prev, f_RS_Prev, f_RL_Prev, ncol=2)
```


```{r}
bac_asvs <- c(bac_RS_asv_ids, bac_RL_asv_ids)
bac_asvs_unique <- unique(bac_asvs)
#RL: 242 Total
#RS: 396 total
#471 unique ASVs

table(bac_RL_asv_ids %in% bac_RS_asv_ids)
#75 "unique" to the RL

table(bac_RS_asv_ids %in% bac_RL_asv_ids)
#229 "unique" to the RS

```



